---
title: "Methods"
output: 
  html_document:
    theme: flatly
    toc: true
    toc_float: true
---
# Introduction:
To predict diabetes risk, we used three models: logistic regression, logistic regression with lasso regularization, and random forests. Importantly, our prediction task is a binary classification problem with no diabetes = 0 and diabetes/pre-diabetes = 1. Since the CDC prepared the data by collapsing pre-diabetes and diabetes into one category, we ran our analysis accordingly using two classifiers. 

Logistic regression serves as our baseline model since it predicts binary outcomes while providing highly interpretable results. To select for the most predictive features, we also included a lasso penalty term that selected for only the most predictive features (drove noisy coefficients to zero). We chose L1 (lasso) regularization over L2 (ridge) regularization to identify only the most predictive features in our model. Since L1 enacts a sum of absolute value penalty, whereas L2 enacts a sum of squares penalty term, most coefficients get minimized to zero, thus allowing for only the most predictive features to estimate model output. This decision to use a Lasso model also made the quiz section of this project easier to build and interpret. Lastly, as a robustness check, we included a random forest model to compare the predictive accuracy of our model across different methods.

First, we imported all necessary libraries for our analysis. 
```{r, message = FALSE}
# Import all libraries
library(tidymodels)
library(pROC)
library(dplyr)

# Import the dataset
data <- read.csv("/Users/brigettemahoney/Desktop/diabetes_binary_5050split_health_indicators_BRFSS2015.csv")

```

As discussed in the data preparation section, we checked for null and missing values. Since this is a government-issued dataset, all entries were already cleaned and prepared for analysis. Additionally, since the dataset contains a perfect 50/50 split between patients with and without diabetes, we did not need to do any data resampling to enforce representative between classes. 

# Logistic Regression:
Since this is a supervised learning task, we first defined "diabetes_binary" as our target variable. We split the data into an 80/20 train test split, since this partition is popular and standard for most predictive tasks. Then, we trained a logistic regression model that estimates the probability that a given individual has diabetes based on a linear combination of the 21 features. Unlike linear regression, logistic regression restricts outputs between 0-1, which makes the interpretation of probabilistic risk much more intuitive. 

```{r, message = FALSE}
# Source: Course Website (adapted code from linear regression)
# make the target variable (diabetes) binary 
data$Diabetes_binary <- as.factor(data$Diabetes_binary)

# Split data into 80/20 train/split
set.seed(123)
splits <- initial_split(data, prop = 0.8, strata = Diabetes_binary)
train_data <- training(splits)
test_data <- testing(splits)

# Fit the logistic regression on ALL features
log_model <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(Diabetes_binary ~ ., data = train_data)

# Predict and compute all probabilities
predictions <- predict(log_model, new_data = test_data)
probs <- predict(log_model, new_data = test_data, type = "prob")

# Compute the Accuracy 
accuracy <- mean(predictions$.pred_class == test_data$Diabetes_binary)
print(paste("Accuracy:", round(accuracy, 3)))

# Compute the AUC
roc_curve <- roc(test_data$Diabetes_binary, probs$.pred_1)
print(paste("AUC:", round(auc(roc_curve), 3)))

# Print all of the significant coefficients
tidy(log_model) %>% filter(p.value < 0.05) %>% arrange(desc(abs(estimate)))

# Save the model for quiz later
saveRDS(log_model, "log_model.rds")

```

As will be discussed in more detail in the results section, our logistic regression model produced an Accuracy of 0.743 and an AUC of 0.824. This means that our model correctly classified 74.3% of all patients in the test set. Additionally, for any randomly chosen person with diabetes, our model classified them as higher risk than a randomly chosen person without diabetes 82.4% of the time. 16 out of 21 features were statistically significant (p<0.05), and a more in-depth analysis is detailed in the results section. 

# Lasso Regression
Since 16/21 features were statistically significant, we wanted to explore if we could achieve similar results using only the most important  features. To do so, we introduced a regularization penalty term to drive most of the coefficients to zero. Since logistic regression is a minimizing objective, introducing a penalty term selects only the most important features. As shown below, we chose a penalty of 0.03 since it produced 9 total coefficients. We believe 9 predictive features is an optimal number to base our interactive quiz off of, since it provides a substantial number of features without being exhaustive. As a result, the 9 most predictive features are: HighBP, GenHlth, HighChol, Age, BMI, HeartDiseaseorAttack, CholCheck, Income, HvyAlcoholConsump.

```{r, message = FALSE}

# Run Lasso Regularization 
lasso_model <- logistic_reg(penalty = 0.03, mixture = 1) %>%
  set_engine("glmnet") %>%
  fit(Diabetes_binary ~ ., data = train_data)

# run the lasso model
tidy(lasso_model) %>%
# only keep non-zero coefficients
  filter(estimate != 0) %>%
  arrange(desc(abs(estimate)))

# save the model 
saveRDS(lasso_model, "lasso_model.rds")

# Get all ROC metrics
lasso_probs_roc <- roc(test_data$Diabetes_binary, predict(lasso_model, new_data = test_data, type = "prob")$.pred_1)
saveRDS(lasso_probs_roc, "lasso_roc.rds")

# compute and save accuracy metrics
lasso_acc <- mean(predict(lasso_model, test_data)$.pred_class == test_data$Diabetes_binary)
saveRDS(lasso_acc, "lasso_acc.rds")

```
# Random Forest Model
As a robustness check, we created a random forest model. Random forests create a sequence of decision trees that can capture non-linear relationships between features and outcomes. As shown in the code below, we fit the random forest model using 500 trees and evaluated their performance using 5-fold cross validation. We chose these model parameters to optimize for the model accuracy / running time tradeoff. Importantly, we used the 9 selected features from the lasso model to fit the model. As a result, the model produced the same AUC as logistic regression (0.824), which therefore validates our results. 

```{r, message = FALSE}

# Source: Course Website

library(doParallel)
registerDoParallel(parallel::detectCores() - 1)

# create a random forest model
rf_model <- 
  rand_forest(trees = 500, mtry = tune(), min_n = tune()) %>% 
  set_engine("ranger") %>% 
  set_mode("classification")

# define the random forest workflow
rf_wflow <- 
  workflow() %>% 
  add_formula(Diabetes_binary ~ .) %>% 
  add_model(rf_model)

# set up cross validation
diabetes_folds <- vfold_cv(train_data, v = 3)

keep_pred <- control_resamples(save_pred = TRUE, save_workflow = TRUE)

rf_grid <- grid_regular(
  mtry(range = c(2, 10)),
  min_n(range = c(2, 20)),
  levels = 3
)

# create a resample object
rf_res <- 
  rf_wflow %>% 
  tune_grid(resamples = diabetes_folds, grid = rf_grid, control = keep_pred)

# examine the metrics across folds
collect_metrics(rf_res)

best_rf <- select_best(rf_res, metric = "roc_auc")

final_rf_wflow <- finalize_workflow(rf_wflow, best_rf)
rf_fit <- final_rf_wflow %>% fit(data = train_data)

final_rf_res <- last_fit(final_rf_wflow, splits)
collect_metrics(final_rf_res)

# AUC
rf_probs <- predict(rf_fit, new_data = test_data, type = "prob")
rf_roc <- roc(test_data$Diabetes_binary, rf_probs$.pred_1)
print(paste("Random Forest AUC:", round(auc(rf_roc), 3)))

# save model
saveRDS(rf_fit, "rf_model.rds")

# save accuracy values
rf_acc <- mean(predict(rf_fit, test_data)$.pred_class == test_data$Diabetes_binary)

```


```{r, message = FALSE}
# save all model metrics
saveRDS(probs, "log_probs.rds")
saveRDS(predict(lasso_model, new_data = test_data, type = "prob"), "lasso_probs.rds")
saveRDS(rf_probs, "rf_probs.rds")
saveRDS(roc_curve, "log_roc.rds")
saveRDS(rf_roc, "rf_roc.rds")
saveRDS(accuracy, "log_acc.rds")
saveRDS(rf_acc, "rf_acc.rds")

```